---
title: "P8106 Final Report"
output: pdf_document
author: "Si Li, Weiwei Qi and Qimin Zhang"
editor_options: 
  chunk_output_type: console
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(rpart.plot)
library(readxl)
```

# Introduction

Breast cancer is considered as one of the most common types of cancer among women all over the world, and machine learning methods for breast cancer classification has been a hot topic for many years. In this report, we want to try multiple classification methods on [Wisconsin Breast Cancer Database](https://www.kaggle.com/roustekbio/breast-cancer-csv?select=breastCancer.csv) from the University of Wisconsin Hospitals, Madison from Dr. William H. Wolberg, and compare their performance to see which one is the best for breast cancer classification.

```{r, include=FALSE}
data = read_csv("data.csv") %>% 
  janitor::clean_names() %>% 
  dplyr::select(-id) %>%
  mutate(
    class = case_when(
      class == 2 ~ 'B',
      class == 4 ~ 'M'
    ) %>% as.factor(),
    bare_nucleoli = na_if(bare_nucleoli, '?'),
    bare_nucleoli = as.integer(bare_nucleoli)
  ) %>%
  mutate_if(is.numeric, ~replace(., is.na(.), median(., na.rm = T)))
```

The dataset contains `r dim(data)[1]` tumor subjects and `r dim(data)[2]-1` predictive variables, and the variable 'class' is the type of this tumor, where 'M' is malignant and 'B' is benign. Explanation for some variables:
'clump_thickness': Thickness of clump, from 1 to 10;
'size_uniformity': Uniformity of cell size, from 1 to 10;
'bland_chromatin': Bland chromatin, from 1 to 10.

The variable 'bare_nucleoli' was recorded in the form of character and NA was denoted by '?'. When cleaning the data, we replace '?' by NA and fill it with median.

# Exploratary Analysis


```{r echo=FALSE, fig.height=4}
data %>%
  keep(is.numeric) %>%  
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(bins = 10) 
```

```{r}
data %>% 
  ggplot(aes(x = class, fill = class)) + 
  geom_bar()
```

```{r}

```


# Models

# Conclusions
